Alone Together: Compositional Reasoning and Inference for Weak Isolation

### 2017-10-26 related-work.tex

\paragraph{Specifying weak isolation.}
Adya~\cite{adyaphd} specifies several weak isolation levels in terms
of \emph{dependency graphs} between transactions, and the kinds of
dependencies that are forbidden in each case. The operational nature
of Adya's specifications make them suitable for runtime monitoring and
anomaly detection~\cite{kemmevldb,feketesigmod08,pssi2011}, whereas
the declarative nature of our specifications make them suitable for
formal reasoning about program behavior. Sivaramakrishnan \emph{et
al.}~\cite{pldi15} specify isolation levels declaratively as trace
well-formedness conditions, but their specifications implicitly assume
a complete trace with only committed transactions, making it difficult
to reason about a program as it builds the trace. Cerone \emph{et
al.}~\cite{gotsmanconcur15} specify isolation levels with atomic
visibility, but their specifications are also for complete traces.
Like ours, all the aforementioned specification frameworks use the
vocabulary introduced in~\cite{burckhardt14}. However, none of them
are equipped with a reasoning framework that can use such
specifications to verify programs under weak isolation.

Recent work by Crooks \emph{et al.}~\cite{CPA+17} also explores the
use of a state-based interpretation of isolation as we do, and share
our position that enables specification of weak isolation that are not
tied to implementation-specific artifacts.  However, they do not
consider verification (manual or automated) of client programs, and it
is not immediately apparent if their specification formalism is
amenable for use within a verification toolchain.

>>>>

\paragraph{Specifying weak isolation.}
Adya~\cite{adyaphd} specifies several weak isolation levels in terms
of \emph{dependency graphs} between transactions, and the kinds of
dependencies that are forbidden in each case. The operational nature
of Adya's specifications make them suitable for runtime monitoring and
anomaly detection~\cite{kemmevldb,feketesigmod08,pssi2011}, whereas
the declarative nature of our specifications make them suitable for
formal reasoning about program behavior. Sivaramakrishnan \emph{et
al.}~\cite{pldi15} specify isolation levels declaratively as trace
well-formedness conditions, but their specifications implicitly assume
a complete trace with only committed transactions, making it difficult
to reason about a program as it builds the trace. Cerone \emph{et
al.}~\cite{gotsmanconcur15} specify isolation levels with atomic
visibility, but their specifications are also for complete traces.
Both the aforementioned specification frameworks use the vocabulary
introduced in~\cite{burckhardt14}. However, none of them are equipped
with a reasoning framework that can use such specifications to verify
programs under weak isolation.

Recent work by Crooks \emph{et al.}~\cite{CPA+17} also explores the
use of a state-based interpretation of isolation as we do, and share
our position that enables specification of weak isolation that are not
tied to implementation-specific artifacts.  However, they do not
consider verification (manual or automated) of client programs, and it
is not immediately apparent if their specification formalism is
amenable for use within a verification toolchain.  ~\cite{WB17}
presents a dynamic analysis for weak isolation that attempts to
discover weak isolation anomalies from SQL log files.  Their solution,
while capable of identifying database attacks due to the use of
incorrect (excessively weak) isolation levels, does not consider how
to verify application correctness, infer proper isolation levels, or
formally reason about the relationship between weak-isolation levels
and application invariants.
